---
title: "Final Project"
author: "Julia Nelson"
date: "12/7/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r set-up}
require(mosaic)
trellis.par.set(theme=col.mosaic())
options(digits = 3)

```
## Question 11.53
```{r q11.53, echo=FALSE}  
  ds = ex11_53cheese
  names(ds)
```
  #### Find mean, median, std, interquartile range
```{r stats} 
  favstats(ds$taste)  
  favstats(ds$acetic) 
  favstats(ds$h2s)    
  favstats(ds$lactic) 
```
Taste:  Mean is 24.5, Median is 20.9 StandardDev is  16.3, InterquartileRange is (13.6,36.7)
Acetic: Mean is 5.50, Median is 5.42 StandardDev is 0.571, InterquartileRange is (5.24,5.88)
H2S:    Mean is 5.94, Median is 5.33 StandardDev is  2.13, InterquartileRange is (3.98,7.57)
Lactic: Mean is 1.44, Median is 1.45 StandardDev is 0.303, InterquartileRange is (1.25,1.67)

  #### Display each distribution's stemplot and Normal Quantile Plot
    ##### Taste:
```{r stem and qqnorm}
      stem(ds$taste)
      qqnorm(ds$taste)
      qqline(ds$taste)
```
      ##### Normality: The taste Distribution is skewed slightly right but is approximately Normal.
    ##### Acetic:
```{r stem nd qqnorm}
      stem(ds$acetic)
      qqnorm(ds$acetic)
      qqline(ds$acetic)
```
      ##### The acetic Distribution is approximately Normal.
    ##### H2S:
```{r stem n qqnorm}
      stem(ds$h2s)
      qqnorm(ds$h2s)
      qqline(ds$h2s)
```
      ##### The H2S Distribution is skewed slightly right but is approximately Normal.
    ##### Lactic:
```{r stem  qqnorm}
      stem(ds$lactic)
      qqnorm(ds$lactic)
      qqline(ds$lactic)
```
      ##### The Lactic Distribution is approximately Normal





## Question 11.54
  
  #### Scatterplot's of each pair of variables
      ##### taste & acetic
```{r pair}
        plot(subset(ds, select = c("taste","acetic")), main = "Scatterplot Taste vs. Acetic")
```
      ##### taste & h2s
```{r pair2}
        plot(subset(ds, select = c("taste","h2s")), main = "Scatterplot Taste vs. H2S")
```
      ##### taste & lactic
```{r pair3}
        plot(subset(ds, select = c("taste","lactic")), main = "Scatterplot Taste vs. Lactic")
```
      ##### acetic & h2s
```{r pair4}
        plot(subset(ds, select = c("acetic","h2s")), main = "Scatterplot Acetic vs. H2S")
```
      ##### acetic & lactic
```{r pair5}
        plot(subset(ds, select = c("acetic","lactic")), main = "Scatterplot Acetic vs. Lactic")
```
      ##### h2s & lactic
```{r pair6}
        plot(subset(ds, select = c("h2s","lactic")), main = "Scatterplot H2S vs. Lactic")
```
  
  #### Correlation & P-value for 0-pop correlation
```{r pair correlation}
    cor.test(ds$taste,ds$acetic)
    cor.test(ds$taste,ds$h2s)
    cor.test(ds$taste,ds$lactic)
    cor.test(ds$acetic,ds$h2s)
    cor.test(ds$acetic,ds$lactic)
    cor.test(ds$h2s,ds$lactic)
```
  #### Outcome Summary:
    ##### Pair          Correlation  P-Values     (closer to 1  is stronger)    
    ##### taste, acetic   0.550     0.002        the weakest linear relationship
    ##### taste, h2s      0.756     0.000001     the strongest positive linear relationship
    ##### taste, lactic   0.704     0.00001      positive linear relationship
    ##### acetic, h2s     0.618     0.0003       positive linear relationship
    ##### acetic, lactic  0.604     0.0004       positive linear relationship
    ##### h2s, lactic     0.645     0.0001       positive linear relationship
      ##### p-values of zero pop correlation are very very small, can reject hypothesis
      ##### that the correlations on pop are not zero



## Question 11.55
    #### simple linear regression 
    #### taste - response variable
    #### acetic - explanatory variable
```{r lm1t}
    acetic = ds$acetic
    taste = ds$taste
    lm1 = lm(taste~acetic, data = ds)
    summary(lm1)  
    anova(lm1)
```
```{r lm1, echo=FALSE} 
    coef(lm1)
    rsquared(lm1)
```
    #### plot of the data 
```{r lm1a}
      plot(acetic, taste, xlab = "Acetic", ylab = "Taste", main = "Plot Taste vs Acetic")
      abline(coef(lm1), lwd =2, lty = 2, col="red") # add the found least-squares regression line
```
    
    #### Summary:
      ##### Equation: $Taste = \ -61.5 + 15.7acetic $
      ##### R-squared =  0.302$
      ##### F-statistic = 12.1
      ##### acetic: coef= 15.7, t-value = 3.48,  std err = 4.5
      ##### intercept: coef= -61.5, t-value = -2.48,  std err = 24.9
      ##### P-value = 0.002
    
  #### The residuals:
    ##### Residuals
```{r lm1b}
        plot(lm1, which =1)
        plot(lm1, which=2)
```
      ##### The Residuals have a relatively Normal distribution
    #### Plot of acetic Residuals vs H2S
```{r lm1c}
      h2s = ds$h2s
      xyplot(residuals(lm1) ~ h2s, type=c("p", "r", "smooth"), main = "Residuals v H2S")
```
      ##### Residuals and H2S are positively associated  with a more random pattern
    #### Plot of acetic Residuals vs Lactic
```{r lm1d}
      lactic = ds$lactic
      xyplot(residuals(lm1) ~ lactic, type=c("p", "r", "smooth"),  main = "Residuals v Lactic")
```
      ##### The residuals seem to be slightly positively associated with lactic but with a random patter
    ##### Residuals Summary:
      ##### Min: -29.64
      ##### 1Q: -7.44
      ##### Median: 2.08
      ##### 3Q: 6.60
      ##### Max: 26.58
      ##### res std error: 13.8 on 28 df
      


## Question 11.56
  #### simple linear regression 
  #### taste - response var
  #### h2s - explanatory var
```{r lm2d}
  taste = ds$taste
  h2s = ds$h2s
  lm2 = lm(taste~h2s, data = ds)
  coef(lm2)
  anova(lm2)
```
```{r lm2t, echo=FALSE}
  rsquared(lm2)
  summary(lm2)
```
  #### plot of the data 
```{r lm2q}
    plot(h2s, taste, xlab = "H2S", ylab = "Taste", main = "Plot Taste vs H2S")
    abline(coef(lm2), lwd =2, lty = 2, col="red") # add the found least-squares regression line
```
 
  #### Summary:
    Equation: $Taste = \ -9.79 + 5.78H2S $
    R-squared = 0.571
    F-statistic = 37.3
    h2s: coef= 5.776, t-value = 6.11,  std err = 0.946
    intercept: coef= -9.787, t-value = -1.64,  std err = 5.958
    P-value = 0.00000137 = < 0.0005
    
  #### The residuals:
    ##### Residuals
```{r lm2}
      plot(lm2, which=1)
      plot(lm2, which=2)
```
      The residuals are of Normal Distribution
    #### Plot of H2S residuals vs acetic
```{r lm2a}
      acetic = ds$acetic
      xyplot(residuals(lm2) ~ acetic, type=c("p", "r", "smooth"), main = "Residuals v Acetic")
```
      There is no real pattern between the residuals and lactic
    #### Plot of H2S residuals vs lactic
```{r lm2s}
      lactic = ds$lactic
      xyplot(residuals(lm2) ~ lactic, type=c("p", "r", "smooth"),  main = "Residuals v Lactic")
```
      There is possibly a  pattern between the residuals and lactic meaning the fit could be better
    #### Residuals Summary:
        Min: -15.43
        1Q: -7.61
        Median: 3.49
        3Q: 6.42
        Max: 25.69
        res std error: 10.8 on 28 df
  

## Question 11.57
  #### simple linear regression 
  #### taste - response var
  #### lactic - explanatory var
```{r lm3d}
  taste = ds$taste
  lactic = ds$lactic
  lm3 = lm(taste~lactic, data = ds)
  summary(lm3)
  anova(lm3)
```
```{r lm3, echo=FALSE}
  coef(lm3)
  rsquared(lm3)
```
    #### plot of the data 
```{r lm3f}
    plot(lactic, taste, xlab = "Lactic", ylab = "Taste", main = "Plot Taste vs Lactic")
    abline(coef(lm3), lwd =2, lty = 2, col="red") # add the found least-squares regression line
```
  
  #### Summary:
        Equation: $Taste = \ -29.9 + 37.7lactic$
        R-squared = 0.496
        F-statistic = 27.5
        lactic: coef= 37.72, t-value = 5.25,  std err = 7.19
        intercept: coef= -29.86, t-value = -2.82,  std err = 10.58
        P-value = 0.000014
    
  #### The residuals:
    ##### Residuals
```{r lm3a}
      plot(lm3, which=1)
      plot(lm3, which =2)
```
        The residuals are of Normal Distribution
    #### Plot of Lactic residuals vs acetic
```{r lm3s}
      acetic = ds$acetic
      xyplot(residuals(lm3) ~ acetic, type=c("p", "r", "smooth"), main = "Residuals v Acetic")
```
        There pattern present between the residuals and acetic
    #### Plot of Lactic residuals vs H2S
```{r lm3w}
      h2s = ds$h2s
      xyplot(residuals(lm3) ~ h2s, type=c("p", "r", "smooth"),  main = "Residuals v H2S")
```
        There pattern present between the residuals and H2S
  #### Residuals Summary:
        Min: -19.94
        1Q: -8.68
        Median: -0.11
        3Q: 9.00
        Max: 27.43
        res std error: 11.7 on 28 df
  


## Question 11.58
  #### Comparing the  3 simple linear regressions
   
  #### The 3 regression equations are:
       $Taste = \ -61.5 + 15.7Acetic$
       $Taste = \ -9.79 + 5.78H2S $
       $Taste = \ -29.9 + 37.7Lactic $
     The intercepts all differ because they are comparing on differently scaled variables
      the intercept changes depensing on when x = 0, and in each equation, x has a different coefficient



## Question 11.59 
    #### multiple linear regression Taste using acetic + h2s
```{r lm4e}
    acetic = ds$acetic
    taste = ds$taste
    h2s = ds$h2s
    lm4 = lm(taste~ acetic + h2s, data = ds)
    plot(lm4, which = 1)
    plot(lm4, which = 2)
    summary(lm4)  
    anova(lm4)
```
```{r lm4t, echo=FALSE}
    coef(lm4)
    rsquared(lm4)
```
    
    #### Summary:
       ##### Equation: Taste =  -26.94 + 3.80Acetic + 5.15H2S
       ##### R-squared = 0.551
       ##### F-statistic = 18.8
       ##### acetic: coef= 3.80, t-value = 0.84,  std err = 4.51, P = 0.406
       ##### h2s: coef= 5.15, t-value = 4.26,  std err = 1.21, P = 0.00022
       ##### intercept: coef= -26.94, t-value = -1.27,  std err = 21.19, P = 0.21454
     Compared to the simple linear regression of Taste~Acetic,  
     where t = 3.48, P = 0.002, the Multiple linear regression including H2S 
     gave Acetic t = 0.84, P= 0.406. 
     Because the two are correlated, not 
     I would prefer the multiple linear regression model however, because 
     it only a slightly  better fit and provides 55.1% of the variation in taste
     versus the simple linear regression's 30.2% variation.

    
    
## Question 11.60
    #### mult linear reg Taste using Lactic + h2s
```{r lm5}
    lactic = ds$lactic
    taste = ds$taste
    h2s = ds$h2s
    lm5 = lm(taste~ h2s + lactic, data = ds)
    summary(lm5)  
    anova(lm5)
    plot(lm5, which=1)
    plot(lm5, which=2)
```
```{r lm5t, echo=FALSE}
    coef(lm5)
    rsquared(lm5)
```
    
  #### Summary :
   Compared to the results to the simple linear regressions using each 
     of these variables alone, it is evident that a better result 
     is obtained by using both predictors in a model. 
     In this model,
       Intercept: coef= -27.59, std Error= 8.98, t-value= -3.07, Pr(>|t|)= 0.0048
       #####H2S: coef= 3.95, std Error= 1.14, t-value= 3.47, Pr(>|t|)= 0.0017
       #####Lactic: coef= 19.89, std Error= 7.96, t-value= 2.50, Pr(>|t|)= 0.0188
     In the individual simple linear regressions:
      #####H2S: coef= 5.776,  std err = 0.946, t-value = 6.11, Pr(>|t|)= 0.0000014
      #####Lactic: coef= 37.72, std err = 7.19, t-value = 5.25, Pr(>|t|)=0.000014
     Comparing the outputs, it is clear that for both variables the multiple 
     linear regression model is better. You can see that the R-squared value (= 0.652) 
     is also higher than the individual regression's (0.571, 0.496) providing more information 
     on the variation of taste. In addition, both have higher P values in the 
     multiple linear regression signifying the coeffients are significantly different than 0.
    
    
    
## Question 11.61 
    #### multiple linear regression - all
```{r lmALLt}
    lactic = ds$lactic
    taste = ds$taste
    h2s = ds$h2s
    acetic = ds$acetic
    lmALL = lm(taste~ acetic+ h2s + lactic, data = ds)
    summary(lmALL)  
    anova(lmALL)
    plot(lmALL, which=1)
    plot(lmALL, which=2)
```
```{r lmALL, echo=FALSE}
    coef(lmALL)
    rsquared(lmALL)
```
    #### Summary:
       Equation: hatTaste =  -28.877 + 0.328Acetic + 3.912H2S + 19.671Lactic
       R-squared = 0.652
       F-statistic = 16.2
       acetic:  t-value = 0.07, P = 0.9420
       h2s:  t-value = 3.13, P = 0.0042
       lactic:  t-value = 2.28, P = 0.0331
       intercept:  t-value = -1.46, P = 0.1554
    #### Residuals:
```{r lmALLa}
      plot(lmALL, which=1)
      plot(lmALL, which =2)
```
       The residuals seem to be Normally distributed 

      #### scatter Plot of  residuals vs acetic
```{r lmALLs}
        xyplot(residuals(lmALL) ~ acetic, type=c("p", "r", "smooth"), main = "Residuals v Acetic")
```
      #### scatter Plot of  residuals vs H2S
```{r lmALLd}
        xyplot(residuals(lmALL) ~ h2s, type=c("p", "r", "smooth"),  main = "Residuals v H2S")
```
      #### scatter Plot of  residuals vs Lactic
```{r lmALLf}
        xyplot(residuals(lmALL) ~ lactic, type=c("p", "r", "smooth"), main = "Residuals v Lactic")
```
       No significant patterns seen in any of the residuals v variables scatterplots
       Comparing the R-squared of lmALL to lm5, they both = 0.652, but because Acetic is very close to 0 with P= 0.9420, 
         making it almost unnecessary to include in the model. The best model to choose would be lm5, with H2S and Lactic, 
         and not lmALL because of the unnecessary information of acetic.
